---
title: "GLMM and GLM Analysis"
author: "Rui Carvalho"
output:
  html_document:
    theme: united
    highlight: tango
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
---

```{r setup, }
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
options(knitr.table.format = "html")

```


# Preparing files

## Libraries 

``` {r, results ='hide'}  
library(BAT)
library(readr)
library(FD) 
library(car)
library(MASS)
library(lme4)
library(here)       # to locate files
library(data.table) # to work with data
library(dplyr)      # to manage data
library(magrittr)   # to use the pipe operator %>% 
library(MuMIn)
library(glmmTMB)
library(bbmle)
```


## Calling results dataset

```{r}
Results2 <- read.csv2(here("results","RESULTS.csv"), header=TRUE, row.names = 1,  stringsAsFactors = T, sep = ",", dec = ".")
str(Results2)
```

## Checking for colinearity 

```{r}
results.variables <- Results2[,5:7] # selecting the variables relevant to the test

numeric.results.variables <- Results2[,5:7] # selecting the variables relevant to the test
cor(numeric.results.variables)  #isto ontem funcionava, não sei o que se passa

```

# Alpha Taxonomic - All Species


```{r}

gm1 <- glmmTMB(TAlphaAll ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = poisson)
options(na.action = "na.fail")
dredge(gm1)
```

# Alpha Taxonomic - Native Species


```{r}

gm2 <- glmmTMB(TAlphaNat ~ Dist_edge_std +Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = poisson)
dredge(gm2)
```

# Alpha Taxonomic - Nind species


```{r}

gm3 <- glmmTMB(TAlphaNInd ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = poisson)
summary(gm3)

```


# Alpha Fuctional - All Species


```{r, }

gm4 <- glmmTMB(FAlphaAll ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = Gamma)
dredge(gm4)
```

# Alpha Functional - Native Species


```{r, }

gm5 <- glmmTMB(FAlphaNat ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = Gamma)
dredge(gm5)
```

# FAILED Alpha Functional - N/Ind Species


``{r, }
gm6 <- glmmTMB(FAlphaNInd ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = Gamma)
dredge(gm6)
``
FAILED - Error in eval(family$initialize, rho) : non-positive values not allowed for the 'gamma' family


# Beta Taxonomic - All Species


```{r, }

gm7 <- glmmTMB(BetaAllTotalVector ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm7)

```


```{r, }
gm8 <- glmmTMB(BetaNatTotalVector ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm8)
```


# Beta Taxonomic - NInd Species


```{r, }

gm9 <- glmmTMB(BetaNIndTotalVector ~ Dist_edge_std +  Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm9)
```



# Beta Functional - All 


```{r, }

gm10 <- glmmTMB(BetaFuncAllTotalVector ~ Dist_edge_std +Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm10)
```

# Beta Functional - Nat 


```{r, }

gm11 <- glmmTMB(BetaFuncNatTotalVector ~ Dist_edge_std +  Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm11)
```


# Beta Functional - NInd 


```{r, }

gm12 <- glmmTMB(BetaFuncNIndTotalVector ~ Dist_edge_std +  Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm10)
```




