---
title: "GLMM and GLM Analysis"
author: "Rui Carvalho"
output: 
  html_document:
    theme: united
    highlight: tango
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
---

```{r setup, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
options(knitr.table.format = "html")

```


# Preparing files

## Libraries 

``` {r, results ='hide'}  
library(BAT)
library(readr)
library(FD) 
library(car)
library(MASS)
library(nlme)
library(lme4)
library(here)       # to locate files
library(data.table) # to work with data
library(dplyr)      # to manage data
library(magrittr)   # to use the pipe operator %>% 
library(MuMIn)
library(glmmTMB)
library(bbmle)
```


## Calling results dataset

```{r}
Results2 <- read.csv2(here("results","RESULTS.csv"), header=TRUE, row.names = 1,  stringsAsFactors = T, sep = ",", dec = ".")
str(Results2)

fam <- read.csv2(here("results","RESULTS copy.csv"), header=TRUE, row.names = 1,  stringsAsFactors = T, sep = ";", dec = ".")
fam <- fam[1,]

```

#Test results and analysis 
When model fits are ranked according to their AIC values, the model with the lowest AIC value being considered the ???best???. Models in which the difference in AIC relative to AICmin is < 2 will be considered also to have substantial support (Burnham, 2002; Burnham and Anderson, 1998).
I will then assess the variables importance, by assessing their respective weights within each models fit. 
By these two criteria, I will then elavborate the discussion on what factors do accet the spiders community. 

## Checking for colinearity 

```{r}
results.variables <- Results2[,5:7] # selecting the variables relevant to the test

numeric.results.variables <- Results2[,5:7] # selecting the variables relevant to the test
cor(numeric.results.variables)  #isto ontem funcionava, não sei o que se passa

```

# Alpha Taxonomic - All Species


```{r}

gm1 <- glmmTMB(TAlphaAll ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = poisson)
options(na.action = "na.fail")   #na.fail is the one that works
dredgm1 <- dredge(gm1)
dredgm1

formula(gm1)

dredge(gm9)
dredgm9 <- dredge(gm9)
subsetgm9 <- subset(dredgm9, dredgm9$delta <= 2)
str(subsetgm9)

subsetgm11 <- filter(dredgm9, dredgm9$delta <= 2)
str(subsetgm11)

bbbb <- getResponseFormula(gm1)
bbbb <- data.frame(bbbb)
write.csv(Results, file = here("results","bbbb.csv"), row.names = TRUE)


str(aaaa)
```

# Alpha Taxonomic - Native Species


```{r}

gm2 <- glmmTMB(TAlphaNat ~ Dist_edge_std +Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = poisson)
dredge(gm2)
```

# Alpha Taxonomic - Nind species


```{r}

gm3 <- glmmTMB(TAlphaNInd ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = poisson)
coeffs(gm3)
modgm3 <- dredge(gm3)
str(modgm3)
coeffs(modgm3)
write.csv(modgm3, file = here("results","modgm3.csv"), row.names = TRUE)
```

gm1$
# Alpha Functional - All Species


```{r, warning=FALSE}

gm4 <- glmmTMB(FAlphaAll ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = Gamma)
dredge(gm4)
```

# Alpha Functional - Native Species


```{r, warning=FALSE}

gm5 <- glmmTMB(FAlphaNat ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = Gamma)
dredge(gm5)
```

# FAILED Alpha Functional - N/Ind Species


``{r, warning=FALSE}
gm6 <- glmmTMB(FAlphaNInd ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = Gamma)
dredge(gm6)
``
FAILED - Error in eval(family$initialize, rho) : non-positive values not allowed for the 'gamma' family


# Beta Taxonomic - All Species


```{r, warning=FALSE}

gm7 <- glmmTMB(BetaAllTotalVector ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm7)

```


## 
```{r, warning=FALSE}
gm7.5  <- glmmTMB(BetaAllTotalVector ~ Dist_trail_std + (1 | ForestID), data = Results2 , family = beta_family)
summary(gm7.5)
```

##
```{r, warning=FALSE}
gm7.6  <- glmmTMB(BetaAllTotalVector ~ Dist_edge_std + (1 | ForestID), data = Results2 , family = beta_family)
summary(gm7.6)
```

##
```{r, warning=FALSE}
gm7.7  <- glmmTMB(BetaAllTotalVector ~ Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
summary(gm7.7)
```














# Beta Taxonomic - Nat Species

```{r, warning=FALSE}
gm8 <- glmmTMB(BetaNatTotalVector ~ Dist_edge_std + Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm8)
```

## Testing specific models

```{r, warning=FALSE}
gm8.6 <- glmmTMB(BetaNatTotalVector ~ Dist_edge_std + Dist_trail_std + (1 | ForestID), data = Results2 , family = beta_family)
summary(gm8.6)
```

```{r, warning=FALSE}
gm8.7 <- glmmTMB(BetaNatTotalVector ~  Dist_trail_std  + (1 | ForestID), data = Results2 , family = beta_family)
summary(gm8.7)
```


# Beta Taxonomic - NInd Species


```{r, warning=FALSE}

gm9 <- glmmTMB(BetaNIndTotalVector ~ Dist_edge_std +  Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm9)
```

```{r, warning=FALSE}

gm9.5 <- glmmTMB(BetaNIndTotalVector ~  Dist_trail_beginning_std +  (1 | ForestID), data = Results2 , family = beta_family)
summary(gm9.5)
```




# Beta Functional - All 


```{r, warning=FALSE}

gm10 <- glmmTMB(BetaFuncAllTotalVector ~ Dist_edge_std +Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm10)
```
## Exploring best ranked models
```{r, warning=FALSE}

gm10.5 <- glmmTMB(BetaFuncAllTotalVector ~ Dist_trail_std +  (1 | ForestID), data = Results2 , family = beta_family)
summary(gm10.5)
```


```{r, warning=FALSE}

gm10.6 <- glmmTMB(BetaFuncAllTotalVector ~ Dist_edge_std +Dist_trail_std + (1 | ForestID), data = Results2 , family = beta_family)
summary(gm10.6)
```
# Beta Functional - Nat 


```{r, warning=FALSE}

gm11 <- glmmTMB(BetaFuncNatTotalVector ~ Dist_edge_std +  Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm11)
```

## Exploring best ranked models
```{r, warning=FALSE}

gm11.5 <- glmmTMB(BetaFuncNatTotalVector ~  Dist_trail_std + (1 | ForestID), data = Results2 , family = beta_family)
summary(gm11.5)
```

```{r, warning=FALSE}

gm11.6 <- glmmTMB(BetaFuncNatTotalVector ~ Dist_edge_std +  Dist_trail_std + (1 | ForestID), data = Results2 , family = beta_family)
summary(gm11.6)
```




# Beta Functional - NInd 


```{r, warning=FALSE}

gm12 <- glmmTMB(BetaFuncNIndTotalVector ~ Dist_edge_std +  Dist_trail_std + Dist_trail_beginning_std + (1 | ForestID), data = Results2 , family = beta_family)
dredge(gm12)
```

## Exploring best ranked models

```{r, warning=FALSE}

gm12.6 <- glmmTMB(BetaFuncNIndTotalVector ~ Dist_edge_std +  Dist_trail_std + (1 | ForestID), data = Results2 , family = beta_family)
summary(gm12.6)
```








