---
title: "workfile.results.analysis"
author: "Rui Carvalho"
date: "1/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }

prop.Talpha.glmm.16 = glmmTMB(prop.Talpha ~ Dist_edge_std +  Dist_trail_beginning_std +(1 | ForestID), data= Results2,family = "beta_family") 
model.output(prop.Talpha.glmm.16) 


```
```{r }

prop.Talpha.glmm.16 = glmmTMB(prop.Talpha ~ Dist_edge_std +  Dist_trail +(1 | ForestID), data= Results2,family = "beta_family") 
model.output(prop.Talpha.glmm.16) 


```




##Variável Abind Nind

```{r }

abund.nind.glmm.14 = glmmTMB(abund.nind ~  Dist_trail_std +(1 | ForestID), data= Results2,family = "nbinom1") 
model.output(abund.nind.glmm.14)  


```

dddddd



```{r }

abund.nind.glmm.13 = glmmTMB(abund.nind ~ Dist_edge_std +  (1 | ForestID), data= Results2,family = "nbinom1") 
model.output(abund.nind.glmm.13)  


```


FAlphaNat.glmm.9 = glmmTMB(FAlphaNat ~    (1 | ForestID), data= Results2,family = "Gamma") 











```{r }

abund.nind.glmm.14 = glmmTMB(abund.nind ~  Dist_trail_std +(1 | ForestID), data= Results2,family = "nbinom1") 
model.output(abund.nind.glmm.14)  


```












```{r }

abund.nind.glmm.14 = glmmTMB(abund.nind ~  Dist_trail_beginning_std + Dist_edge_std + (1 | ForestID), data= Results2,family = "nbinom1") 
model.output(abund.nind.glmm.14)  


```



```{r }

FAlphaNat.glmm.9 = glmmTMB(FAlphaNat ~  Dist_trail_std  (1 | ForestID), data= Results2,family = "Gamma") 
model.output(FAlphaNat.glmm.9)  


```



### Beta Diversity analysis

```{r }

all.tax.btotal.glmm.20 = glmmTMB(all.tax.btotal ~   Dist_trail_std +(1 | ForestID), data= Results2,family = "beta_family") 
model.output(all.tax.btotal.glmm.20)   


simul.all.tax.btotal.glmm.20 <- simulateResiduals(fittedModel = all.tax.btotal.glmm.20, n=250)
plot(simul.all.tax.btotal.glmm.20,quantreg = TRUE) 
testResiduals(simul.all.tax.btotal.glmm.20)
```




Burrow_FS_1 <- glmmTMB(Burrowing_T ~ Treatment*Weight_C + DailySeen_C + (1|Colony) +(1|AnimalID), data = CS_FS_Burrow , family=beta_family())

I simulate the residuals and run the diagnostic plot

Simul_FS_Burrow <- simulateResiduals(fittedModel = Burrow_FS_1, n = 250) 
plot(Simul_FS_Burrow,quantreg = TRUE) testResiduals(Simul_FS_Burrow)

